Intentional release of native species undermines ecological stability

### 2022-12-23 rmd/manuscript.Rmd
edit: manuscript correct typo

When small carrying capacity was combined with a low to modest intrinsic growth rate ($K=$ `r min(list_param[[2]]$k)`, $r_1 \le 2.5$), our model predicted a destabilizing effect of intentional release on ecological communities, as illustrated by increased CV with increasing numbers of releases (**Figure** 1)**A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure** 1)**A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, as reported in a previous theoretical study [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure** 1)**A**). Combined, the total community density decreased more sharply than individual species groups (**Figure** 1)**A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure** 1)**A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (**Figures S3 -- S6**), but except for the scenario in which the enhanced species exhibits chaos ($r_1=$ `r max(list_param[[2]]$r1)`).

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with the intensive release (**Figure** 1)**B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure** 2; see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob) %>% round(2), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures** 1)**B and** 2), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures** 1)**B and** 2**;** see **Tables S6 -- S8** for full statistics). In the meantime, the SDs had vague relationships with the intentional release (**Figures** 1)**B and** 2).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimates of interspecific competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure** 3)), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, competition between masu salmon and other salmonid species is well documented, and its form can be interference or exploitative [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004; @moritaImpactExoticTrout2004]. Diets of salmonid species are innately overlapped, and they interfere with each other to claim profitable foraging spots for drift insects [@hasegawaComparisonCompetitiveAbility2004] or cruise over a large area in search of prey [@hasegawaEffectFlowRegime2010]. Exploitation competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014] that many benthic fish species prey on [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmon are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

>>>>

When small carrying capacity was combined with a low to modest intrinsic growth rate ($K=$ `r min(list_param[[2]]$k)`, $r_1 \le 2.5$), our model predicted a destabilizing effect of intentional release on ecological communities, as illustrated by increased CV with increasing numbers of releases (**Figure** 1**A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure** 1**A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals, as reported in a previous theoretical study [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure** 1**A**). Combined, the total community density decreased more sharply than individual species groups (**Figure** 1**A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure** 1**A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (**Figures S3 -- S6**), but except for the scenario in which the enhanced species exhibits chaos ($r_1=$ `r max(list_param[[2]]$r1)`).

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) at sites with the intensive release (**Figure** 1**B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of "effective" releases (**Figure** 2; see **Methods** for the Bayesian estimation of effective release), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob) %>% round(2), 2)` (**Table S6**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures** 1**B and** 2), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures** 1**B and** 2**;** see **Tables S6 -- S8** for full statistics). In the meantime, the SDs had vague relationships with the intentional release (**Figures** 1**B and** 2).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data (nine sites with few missing observations and no intentional release; **Methods**). The parameter estimates of this model are directly comparable to those used in the theoretical Ricker model. The estimates of interspecific competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure** 3), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, competition between masu salmon and other salmonid species is well documented, and its form can be interference or exploitative [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004; @moritaImpactExoticTrout2004]. Diets of salmonid species are innately overlapped, and they interfere with each other to claim profitable foraging spots for drift insects [@hasegawaComparisonCompetitiveAbility2004] or cruise over a large area in search of prey [@hasegawaEffectFlowRegime2010]. Exploitation competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014] that many benthic fish species prey on [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmon are larger and more aggressive than wild individuals of a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].
